### 2015-05-14 intro.tex
POR
hacks in introduction

A major reason behind these concurrency bugs is that device drivers operate inside a complicated and hostile environment~\cite{corbet2005linux}. For example the Linux kernel is able to invoke its drivers from multiple concurrent threads, a hardware device is able to issue interrupt requests that can cause running processes to block and switch execution context, and the user may decide to remove a device while there is still some running operation. All these scenarios can easily cause \emph{data races} if proper synchronization mechanisms (such as locks) are not in place to control concurrent access to shared resources.

@@

In this paper, we present \whoop, a new fully automatic approach that (i) statically analyzes device drivers written in the C language for data races and (ii) exploits the race-related results from the symbolic analysis to accelerate bug-finding. \whoop is empowered by \emph{symbolic pairwise lockset analysis}, a novel technique for verifying the absence of data races in device drivers. We attempt to prove a driver race-free by: (i) deriving a sound \emph{sequential} program that \emph{over-approximates} the originally concurrent program, (ii) instrumenting it to record \emph{locksets}, and (iii) using the locksets to assert that all accesses to the same shared resource are protected by at least one common lock. Reducing analysis to reasoning over a sequential program avoids the need to enumerate thread interleavings, and allows reuse of existing successful sequential verification techniques.

We then show that by exploiting the race-freedom guarantees provided by our symbolic analysis we can achieve a sound partial-order reduction that significantly accelerates \corral~\cite{lal2012corral}, a state-of-the-art bug-finder used by Microsoft to analyze device drivers, as well as other concurrent programs. Using \whoop and \corral we analyzed \sizeOfBenchmarks drivers available in the 4.0 Linux kernel distribution.  We find that by combining \whoop and \corral we achieve concurrency analysis speedups in the range of 1.5--10$\times$ for most of our benchmarks, compared with using \corral in isolation.  For two drivers, we observe even greater speedups of \ADComment{first speedup} and \ADComment{second speedup}.

@@

To summarize, our contributions are as follows:
\begin{itemize}
\item We propose symbolic pairwise lockset analysis, a sound and scalable technique for automatically verifying the absence of data races in device drivers.
\item We present \whoop, a tool that leverages our approach for analyzing data races in device drivers.
\item We show that we can achieve a sound partial-order reduction using our technique to accelerate \corral, a state-of-the-art bug-finder.
\item We analyze \sizeOfBenchmarks Linux drivers, showing that \whoop is efficient at race checking and accelerating \corral.
\end{itemize}

>>>>

A major reason behind these concurrency bugs is that drivers operate inside a complicated and hostile environment~\cite{corbet2005linux}. For example the Linux kernel can invoke its drivers from multiple concurrent threads, a hardware device can issue interrupt requests that cause running processes to block and switch execution context, and the user may decide to remove a device (hot-plugging) while some operation is still running. All these scenarios can easily cause \emph{data races} if no proper synchronization mechanisms (e.g.\ locks) are in place to control concurrent access to shared resources.

@@

In this paper, we present \whoop, a new fully automatic approach that statically analyzes device drivers written in C for data races. \whoop is empowered by \emph{symbolic pairwise lockset analysis}, a novel technique for verifying the absence of data races in device drivers. We attempt to prove a driver race-free by: (i) deriving a sound \emph{sequential} program that \emph{over-approximates} the originally concurrent program; (ii) instrumenting it to record \emph{locksets}; and (iii) using the locksets to assert that all accesses to the same shared resource are \emph{consistently} protected by at least one common lock. Reducing analysis to reasoning over a sequential program avoids the need to enumerate thread interleavings, and allows reuse of existing successful sequential verification techniques.

We then show that by exploiting the race-freedom guarantees provided by our symbolic analysis we can achieve a sound partial-order reduction that significantly accelerates \corral~\cite{lal2012corral}, a state-of-the-art precise bug-finder used by Microsoft to analyze drivers, as well as other concurrent programs. Using \whoop and \corral we analyzed \sizeOfBenchmarks drivers from the 4.0 Linux kernel distribution.  We find that by combining \whoop and \corral we achieve concurrency analysis speedups in the range of 1.5--10$\times$ for most of our benchmarks, compared with using \corral in isolation.  For two drivers, we observe even greater speedups of 12$\times$ and 20$\times$.

@@

To summarize, our contributions are as follows:
\begin{itemize}
\item We propose symbolic pairwise lockset analysis, a sound and scalable technique for automatically verifying the absence of data races in device drivers.
\item We present \whoop, a tool that leverages our approach for analyzing data races in device drivers.
\item We show that we can achieve a sound partial-order reduction using our technique to accelerate \corral, an industrial-strength bug-finder.
\item We analyze \sizeOfBenchmarks Linux drivers, showing that \whoop is efficient at race-checking and accelerating \corral.
\end{itemize}
